---
title: "RANN"
author: "Seonjoo Lee"
date: "`r Sys.Date()`"
output: 
  html_document:
    code_folding: hide
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,message=FALSE, warning=FALSE)
rm(list=ls())
library(psych)
library(ggplot2)
library(reshape2)
library(gridExtra)
library(parallel)
library(lavaan)
library(smm)
library(parallel)
library(dplyr)
library(knitr)
library(ppcor)
library(kableExtra)
load('thick.dat')
```

# Mediation Analysis for RANN 

## Model 1. Mediation: Age -> Mean thickness -> Global cognition
```{r, message=FALSE,warning=FALSE}
B=1000
########################
# 1) Mediation: Age -> Mean thickness -> Global cognition
model <- ' # direct effect
             Global ~ c*Age
          # mediator
          Mean.Thickness ~ a*Age
          Global ~ b*Mean.Thickness
          # indirect effect (a*b)
          ab := a*b
          # total effect
          total := c + (a*b)
'
set.seed(20200512)
fit <- sem(model, data = data.frame(thick), se="boot",bootstrap = B)

kable(parameterEstimates(fit,standardized=TRUE), digits=3)## significant mediation)

a<-ggplot(thick,aes(x=Age,y=Global))+geom_point()+stat_smooth(size=2,method="lm")
b<-ggplot(thick,aes(x=Age,y=Mean.Thickness))+geom_point()+stat_smooth(size=2,method="lm")
c<-ggplot(thick,aes(x=Mean.Thickness,y=Global))+geom_point()+stat_smooth(size=2,method="lm")
```

```{r, fig.width=8, fig.height=4, message=FALSE,warning=FALSE}
grid.arrange(a,b,c, ncol=3, nrow=1)  
png(paste('Figure4_golbal_cog.png',sep=""),width=2700,height=900,res=300)
grid.arrange(a,b,c, ncol=3, nrow=1)  
dev.off()
```

## Model2. p-value based screening

* For the threshold, p-value of 0.2, 0.1, 0.05 and multiple compariation correction controlling for false discovery rate (B-H) will be used. 

#### (a) p<0.2
```{r, message=FALSE,warning=FALSE}
#patha
x = scale(thick$Age)
m = scale(thick[,c(1,12:332)])
y = scale(thick$Global)

pthresh=0.2
patha=which(unlist(lapply(1:ncol(m), function(j)cor.test(x,m[,j])$p.value))<pthresh)
pathb=which(unlist(lapply(1:ncol(m), function(j)pcor.test(y,m[,j],x)$p.value))<pthresh)
pathb[pathb %in% patha]

colnames(m)[pathb[pathb %in% patha]]
```

#### (b) p<0.1 and p<0.05 (results were same)
```{r, message=FALSE,warning=FALSE}
pthresh=0.1
patha=which(unlist(lapply(1:ncol(m), function(j)cor.test(x,m[,j])$p.value))<pthresh)
pathb=which(unlist(lapply(1:ncol(m), function(j)pcor.test(y,m[,j],x)$p.value))<pthresh)
colnames(m)[pathb[pathb %in% patha]]

model.screen <- ' # direct effect
             Global ~ c*Age
          # mediator
          Mean.Thickness ~ a1*Age
          PC002 ~ a2*Age
          PC003 ~ a3*Age
          PC004 ~ a4*Age
          PC012 ~ a5*Age
          Global ~ b1*Mean.Thickness + b2*PC002+b3*PC003+b4*PC004+b5*PC012

          # indirect effect (a*b)
          ab1 := a1*b1
          ab2 := a2*b2
          ab3 := a3*b3
          ab4 := a4*b4
          ab5 := a5*b5
          # total effect
          total := c + ab1 + ab2 + ab3 + ab4 + ab5
'
set.seed(20200512)

fit <- sem(model.screen, data = data.frame(Age=x, Global=y, m), se="boot",bootstrap = B)
kable(parameterEstimates(fit,standardized=TRUE),digits=3) ## significant mediation)

pthresh=0.05
patha=which(unlist(lapply(1:ncol(m), function(j)cor.test(x,m[,j])$p.value))<pthresh)
pathb=which(unlist(lapply(1:ncol(m), function(j)pcor.test(y,m[,j],x)$p.value))<pthresh)
colnames(m)[pathb[pathb %in% patha]]
```


#### (c) Multiple comparison correction (B-H)

* Nothing was selected.
```{r, message=FALSE,warning=FALSE}
patha=which(p.adjust(unlist(lapply(1:ncol(m), function(j)cor.test(x,m[,j])$p.value)),method='BH')<pthresh)
pathb=which(p.adjust(unlist(lapply(1:ncol(m), function(j)pcor.test(y,m[,j],x)$p.value)),method='BH')<pthresh)
colnames(m)[pathb[pathb %in% patha]]
```


## Method 3. sparse mediation 

### Cross-validation
```{r, message=FALSE,warning=FALSE}
system.time(
fit.cv<-cv.sparse.mediation.sgrplasso(x,m,y,tol=0.0001,K=5,max.iter=50,
                                     lambda1 = exp(seq(-1,-5, length=10)),lambda2=exp(c(-1,1)),
                                     alpha=seq(0.5,0.95, length=4),multicore=10,
                                     group.penalty.factor=rep(1,ncol(m)+1),
                                     penalty.factor=rep(1, ncol(m)*2+1),
                                     seednum=20200615,non.zeros.stop = 200))

cat(paste('lambda1=',fit.cv$cv.lambda1,
'lambda2=',fit.cv$cv.lambda2,
'alpha=',fit.cv$cv.alpha))

fit<-sparse.mediation.grplasso(x,m,y,tol=0.0001,max.iter=50,
                               lambda1 =fit.cv$cv.lambda1,
                               lambda2=fit.cv$cv.lambda2,
                               alpha=fit.cv$cv.alpha,
                               group.penalty.factor=rep(1,ncol(m)+1),
                               penalty.factor=rep(1, ncol(m)*2+1),verbose=FALSE)

cat(paste('Number of mediators:',fit$nmed))
colnames(m)[which(fit$medest!=0)]
```

### Examine stability of CV

* evaluate whether the tuning parameters were selected stably.

```{r,warning=FALSE,message=FALSE}
z.cveval<-lapply( (1:10)*20+20200512,
                 function(num){
                   fit<-c()
                   try(fit.cv<-cv.sparse.mediation.grplasso(x,m,y,tol=0.0001,K=5,max.iter=50,
                                                        lambda1 = exp(seq(-1,-5,length=10)),
                                                        lambda2=exp(c(-1,1)),
                                                        group.penalty.factor=rep(1,ncol(m)+1),
                                                        penalty.factor=rep(1, ncol(m)*2+1),
                                                        alpha=seq(0.5,0.95,length=4),multicore=5,
                                                        seednum=num,
                                                        non.zeros.stop = 150))
                   try(fit<-sparse.mediation.grplasso(x,m,y,tol=0.0001,max.iter=50,
                                                    lambda1 =fit.cv$cv.lambda1,
                                                    lambda2=fit.cv$cv.lambda2,
                                                    alpha=fit.cv$cv.alpha,verbose=FALSE))
                   return(fit)
                   })

par(mfrow=c(2,2))
plot(unlist(lapply(z.cveval, function(zz)zz$lambda1)),main='lambda1',ylim=c(exp(-5),exp(-1)))
plot(unlist(lapply(z.cveval, function(zz)zz$lambda2)),main='lambda2')
plot(unlist(lapply(z.cveval, function(zz)zz$alpha)),main='alpha',ylim=c(0.5,0.95))
plot(unlist(lapply(z.cveval, function(zz)zz$nmed)),main='Number of Mediator')


```

### To compute 95% CI of the final model

```{r,warning=FALSE,message=FALSE}
B=1000
which(fit$medest!=0)

cor.test(thick$Global, thick$Age)

oneunit<-function(seednum){
  set.seed(seednum*2 + 20200605)
  indx=sample(length(x),length(x),replace=TRUE)
  refit<-c()
  try(refit<-sparse.mediation.grplasso(x[indx],as.matrix(m[indx,fit$medest!=0]),y[indx],
                                   tol=0.0001,max.iter=50,
                                   lambda1 =0,lambda2=0,alpha=0,verbose=TRUE))
  returndat<-c()
  try(returndat<-data.frame(c=refit$c, med=t(refit$medest), a=t(refit$hata), b=t(refit$hatb)))
  return(returndat)
}

system.time(z.postinf <-data.frame(do.call(rbind,mclapply( (1:B)*2 +20200512, function(num){oneunit(seednum=num)}, mc.cores=16))) )
z.postinf = z.postinf %>% mutate(total=c + med.1 +  med.2 +  med.3 +  med.4) 

re.dat<-data.frame(t(apply(z.postinf,2, function(xx)c(mean(xx),
                                 quantile(xx, probs=c(0.025,0.975))))))
names(re.dat)<-c('Estimate','LCI95','UCI95')

kable(round(re.dat,4)) %>%
  kable_styling(bootstrap_options = "bordered", full_width = F)
```

### Evaluate the importance of each mediator:

* Resampling and conduct 

```{r bootstrap, message=FALSE,warning=FALSE}
B=100
run_smediation<-function(X,M,Y,K=5, 
                         lambda1 = exp(seq(-1,-5,length=10)),
                         lambda2=exp(c(-1,1)),
                         alpha=seq(0.5,0.95,length=4)){
		fit.cv=cv.sparse.mediation.sgrplasso(X,M,Y,tol=0.0001,K=5,max.iter=50,
                                         lambda1 =lambda1,lambda2=lambda2,
                                         alpha=alpha,multicore=1,
		                                     group.penalty.factor=rep(1,ncol(M)+1),
		                                     penalty.factor=rep(1, ncol(M)*2+1),
		                                     non.zeros.stop = 200)
		#twoway mediation
		finalfit<-sparse.mediation.grplasso(X,M,Y,tol=0.0001,max.iter=50,
                                                    lambda1 =fit.cv$cv.lambda1,
                                                    lambda2=fit.cv$cv.lambda2,
                                                    alpha=fit.cv$cv.alpha)
		return(finalfit)
}
### 95% CI

run_ci<-function(seed,X,M,Y){
	set.seed(seed*2 + 20200606)
	n = nrow(X)
	indx1 = sample(n,n,replace=TRUE)
#	print(indx1)
 	return(run_smediation(X[indx1,],M[indx1,],Y[indx1,]))
}


#system.time(oneinstant<-run_smediation(X=as.matrix(thick$Age),M=as.matrix(thick[,c(1,12:loc)]),Y=as.matrix(thick$Global)))
system.time(z<-mclapply(1:B, function(seed){run_ci(seed=seed, X=as.matrix(x),M=m,Y=as.matrix(y))},mc.cores=8))

#print(a)
#save.image('rann_20200615_with_meanthickness.RData')
```

```{r,fig.width=6, fig.height=5, message=FALSE,warning=FALSE}

#load('rann_20200615_with_meanthickness.RData')
selected=(B-apply(do.call(cbind, lapply(z, function(x)x$medest)),1,function(x)sum(x==0)))/B

data.frame(varname=names(thick)[c(1,12:332)], selected=selected) %>%
  arrange(-selected) %>% filter(selected>0.5)

library(ggplot2)
library(ggthemes)
indx=sort.int(selected,decreasing=TRUE, index.return=TRUE)$ix
aa=ggplot(data.frame(select=selected[indx]*100, names=factor(names(thick)[c(1,12:332)][indx],levels=names(thick)[c(1,12:332)][indx]))[1:20,], aes(x=names,y=select))+geom_bar(stat='identity')+theme(axis.text.x = element_text(angle = 90, hjust = 1),plot.title = element_text(hjust = 0.5)) + ggtitle(paste('% of selection out of',B,'bootstrapping: top 20 mediators')) + ylab('% selection')+xlab('')
aa

png('Figure5_selected_20200615.png',width=2100,height=1500,res=300)
aa
dev.off()
```
